\plaintitle

### 2016-12-17 drafts/labor/cases.tex
piecework_lit drafting

\topic{Research in crowdsourcing has spent the better part of a decade
exploring how to grow the limits of crowdsourcing and
find the boundaries of crowd work and microtasks.}
This has largely involved
  identifying challenges to this form of labor,    overcoming them through novel designs of work--flows and processes, and
  repeating the process~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
The question that has emerged
among these researchers and
through the work that they have produced then
has been driving at \textit{whether} there are limits to crowdsourcing
(and, if so, what factors determine those limits).
Through this lens, we can point to
a number of contributions to the field that have extended the boundaries of crowd work.

The exploration of crowdsourcing's potential and limits has principally looked at
manipulating and extending along three dimensions:
\begin{inlinelist}
  \item \nameref{sec:complexity},
  \item \nameref{sec:decomposition}, and
  \item \nameref{sec:relationships}.
\end{inlinelist}
We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
decomposed,  contextually abstracted, and
made more resilient to attrition of various forms.
We'll also point to corresponding piecework literature addressing these aspects.
\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
Finally, we'll discuss how these elements will serve
to constrain the upper and lower bounds of crowdsourcing as it relates
to the question of the furthest limits of crowdsourcing.

>>>>

{Research in crowdsourcing has spent the better part of a decade
exploring how to grow the limits of crowdsourcing,
finding the boundaries of crowd work and microtasks.}
This has largely involved iteratively
  identifying challenges to increasing complexity and
  overcoming them through novel designs of work--flows and processes~\cite[e.g.][]{bernsteinSoylent,foundry,crowdForgeKittur}.
The question that has emerged then
has been \textit{whether} there are limits to crowdsourcing
--- and if so, what factors determine them.
Through this lens,
a number of contributions to the field have pushed the boundaries of crowd work.

The exploration of crowdsourcing's potential and limits has principally looked at
manipulating and extending along three dimensions:
\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
  \item \namerefl{sec:complexity}
  \item \namerefl{sec:decomposition}
  \item \namerefl{sec:relationships}.
\end{Numberlist}
We'll explore these aspects of crowdsourcing,  discussing the extents to which work can be
decomposed,  contextually abstracted, and
made more resilient to attrition of various forms.
We'll also point to corresponding piecework literature addressing these aspects.
\msb{Our goal will be to use that literature to inform answers to the questions about crowd work (or something like that)}
Finally, we'll discuss how these elements will serve
to constrain the upper and lower bounds of crowdsourcing as it relates
to the question of the furthest limits of crowdsourcing.

### 2016-12-17 drafts/labor/introduction.tex
piecework_lit drafting

\section{Introduction}\label{sec:introduction}
\topic{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
the constituent parts of larger tasks.}
This framing of work into de--contextualized, modular blocks
enables computation to hire workers at scale through open calls on the internet~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in the work whenever their schedules allow,
with little to no awareness of the broader context of the work, and 
with (often) fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{inlinelist}
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work; and
\item \textit{gig work}, typically involving platforms for one--off jobs, like driving, courier services, or administrative support.~
\end{inlinelist}
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
like AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
(e.g., \cite{bernsteinSoylent,vizwiz,paolacci2010running}).

@@

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads:
\begin{enumerate}
  \item \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}[label=(\alph*)]
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364};
  \item \nameref{sec:decomposition}?
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}; and%
  \item \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}
\end{enumerate}
This research has largely sought to answer these questions by examining the present on--demand work phenomenon.
So far, it has not offered a framing for holistically explaining
the developments in worker processes that researchers have developed, or
the emergent phenomena in social environments;
nor has any research, to our knowledge,
gone as far as directly predict future developments.

@@

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by looking for patterns of behavior that the corresponding literature predicts
on this basis, we can
\begin{inlinelist}
  \item make sense of the phenomena so far as part of a much larger series of interrelated events;
  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today;
        and finally,
  \item to the extent that history repeats itself, offer predictions of what on--demand work researchers,
        and workers themselves,
        might expect to see on the horizon.
\end{inlinelist}
For example, we will draw on the piecework literature studying task decomposition, which was historically limited in scope by technological limits in measurement and instrumentation, and leverage that understanding to suggest how modern web technology reconfigures this measurement limit for on--demand crowd work.

>>>>

\section{Introduction}\label{sec:introduction}
{The past decade has seen a flourishing of \textit{on--demand work},
largely driven by the reformulation of work as
constituent parts of larger tasks.}
This framing of work into modular blocks
has enabled the computational hiring \& management of workers at scale~\cite{howe2008crowdsourcing,Bigham2014,crowdworkFuture}.
Distributed paid participants then engage in work whenever their schedules allow,
with little to no awareness of the broader context of the work, and
often with fleeting identities and associations~\cite{martin2014being,uberAlgorithm}.
In this paper, we use the term on--demand work to join a pair of related phenomena:
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
\item \textit{crowd work}, on platforms such as Amazon Mechanical Turk (AMT) and other sites of (predominantly) information work
\item \textit{gig work}, often as platforms for one--off jobs, like driving, courier services, or administrative support.
\end{numberlist}
The realization that complex goals can be accomplished by directing and managing crowds of workers spurred industry to flock to sites of labor
like AMT to explore the limits of this distributed, on--demand workforce.
Researchers have also taken to the space in earnest,
developing systems and designs that enable new forms of production
\cite[e.g.][]{bernsteinSoylent,vizwiz,paolacci2010running}.

@@

\topic{This body of research has broadly sought to answer one central question:
What does the future hold for on--demand work and those who do it?}
Researchers have offered insights on this question along three major threads.
\begin{Numberlist}[itemjoin*={.~And~},itemjoin={.~}]
  \item \nameref{sec:complexity}?
        Specifically,
        \begin{inlinelist}[label=(\alph*)]
          \item How complex are the goals that crowd work can accomplish?, and
          \item What kinds of goals and industries may eventually utilize it?
        \end{inlinelist}
        \cite{foundry,suzukiAtelier,KimStoria,yuanAlmost,YuEncouragingOutside,
              Nebeling:2016:WCW:2858036.2858169,
              Hahn:2016:KAB:2858036.2858364}
  \item \nameref{sec:decomposition}
        \cite{embracingErrorKrishna,bernsteinSoylent,sensitiveTasks,
              LykourentzouPersonalityMatters,KucherbaevReLauncher,
              Law:2016:CKC:2858036.2858144,Cai:2016:CRI:2858036.2858237,
              Chang:2016:ACC:2858036.2858411,Newell:2016:OMA:2858036.2858490}
  \item \nameref{sec:relationships}?
        \cite{turkopticon,storiesIraniSilberman,dynamo,crowdcollab,
              whyWouldAnyoneBrewer,takingAHITMcInnis}.
\end{Numberlist}

This research has largely sought to answer these questions by examining extant on--demand work phenomena.
So far, it has not offered a framing for holistically explaining
the developments in worker processes that researchers have developed, or
the emergent phenomena in social environments;
nor has any research
% , to our knowledge,
gone so far as to directly predict future developments.

@@

More concretely, by looking at on--demand work as
an instantiation (or even a continuation) of piecework,
and by looking for patterns that the corresponding literature predicts
on this basis, we can:
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
  \item make sense of past events as part of a much larger series of an interrelated phenomenon
  \item reflect on differences in the factors that impacted piecework historically and impact on--demand work today
  \item to some extent, offer predictions of what on--demand work researchers,
        and workers themselves,
        might expect to see on the horizon.
\end{numberlist}
For example, we will draw on the piecework literature studying task decomposition,
which was historically limited by technological limits in measurement and instrumentation, and
leverage that understanding to suggest how modern technology affects this mechanism in on--demand work.

### 2016-12-17 drafts/labor/piecework_lit.tex
piecework_lit drafting

\section{A Review of Piecework}

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
  \item define ``piecework'' as researchers in the topic understood it;
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time;
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

\subsection{What is piecework?}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While ``piecework'' has had multiple definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We will follow this history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

In summary, piecework:
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models;
  \item afforded workers freedom in when and how much to work;
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
\begin{Numberlist}
\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the workflow~\cite{noronha2011platemate,bernsteinSoylent}.
\end{Numberlist}
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we'll discuss the history of piecework at a high level
to contextualize the observations and conclusions we later make,
rather than provide a comprehensive historical account.}
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework;
as close to a \textit{thick description} of piecework
as we can afford~\cite{geertz1973interpretation}.

@@

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}[itemjoin*={~and~},itemjoin={~}]
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance;
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries;
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}[itemjoin*={~and~},itemjoin={~}]
  \item railroad workers across the United States;
  \item Airy's employment of \textit{human computers};
  \item domestic and farm work (in particular, the ``match--girls'').

\end{numberlist}

@@

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}

\subsubsection{Railroad and other Industrial Workers}
\topic{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's parallel, but coincidental, eponymously named \textit{Fordism}),
piecework found a new audience in the early and mid--\nth{20} century in industrial work.}
Scientific management argued that it was possible to measure work at
unprecedented resolution and precision~\cite{taylor1914principles,towardsGlobalFordism}.
As \citeauthor{Brown01041990} points out,
piecework most greatly benefits the instrumented measurement of workers,
but in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies'' \cite{nadworny1955scientific}
as rudimentary guides on per--task compensation.
This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
railroad maintenance, automobile manufacturing, and later weapons manufacturing.
\ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

@@

\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

@@

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
In 1912, national coal miners effected an individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Women in the garment industry in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
% and it was this framing on work that
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor himself, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy, in the form of George Airy's ``human computers''.}
% {\citeauthor{grier2013computers}~\cite{grier2013computers} gives one of the earliest accounts of
% piecework, found in George Airy's creation of
% the British Nautical Almanac.}
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
\ali{also talk about the challenge/competition Airy was in with other astronomers,
and the approach of verifying worker results}

\topic{Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.}
For one thing,
Airy had set a policy of firing his computers once they reached age 23.
\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
\begin{Numberlist}
\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
\item it limited workers' ability to organize
      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
\end{Numberlist}}

\onlyinsubfile{
  \printbibliography
  }

>>>>

\onlyinsubfile{\clearpage}
\section{A Review of Piecework}

\topic{The HCI community has used the term ``piecework'' to describe
myriad instantiations of on--demand labor,
but this reference has generally been offered in passing.}
As this paper principally traces a relationship between
the historical piecework and the contemporary crowd work
(or on--demand labor more generally),
this casual familiarity with piecework may prove insufficient.
We'll more carefully discuss piecework in this section in order
to inform the the rest of the argument.
Specifically, we will
\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
  \item define ``piecework'' as researchers in the topic understood it
  \item trace the rise and fall of piecework at a very high level,
        identifying key figures and ideas during this time
\end{inlinelist}
This section is not intended to be comprehensive: instead, it sets up the scaffolding necessary for our later investigations of crowd work's three questions: complexity limits, task decomposition, and worker relationships.

% \onlyinsubfile{\clearpage}
\subsection{What is piecework?}
\topic{Aligning on--demand work with piecework requires an understanding of what piecework is.}
While ``piecework'' has had multiple definitions over time,
we can trace a constellation of characteristics that recur throughout the literature.
We will follow this history of research, collecting
descriptions,
examples, and
provided definitions of piecework, trying
to trace the outline of a working understanding of
\textit{what piecework is}.

@@

In summary, piecework:
\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
  \item paid workers for quantity of work done, rather than time done,
        but occasionally mixed the two payment models
  \item afforded workers freedom in when and how much to work
  \item structured tasks such that people who didn't have the training
        to engage in the traditional labor force could still participate.
\end{inlinelist}

Viewing crowd work as a modern instantiation of piecework is relatively straightforward by this definition.
\begin{Numberlist}
\item platforms such as Mechanical Turk, Uber and TaskRabbit pay by the task, though some such as Upwork do offer hourly rates as well.
\item workers are attracted to these platforms by the freedom they offer to pick the time and place of work~\cite{martin2014being,whyWouldAnyoneBrewer}.
\item system developers as on Mechanical Turk typically assume no professional skills in transcription or other areas, and attempt to build that expertise into the work flow~\cite{noronha2011platemate,bernsteinSoylent}.
\end{Numberlist}
Given this alignment, many of the same properties of piecework historically will apply to on--demand work as well. In the next section, we perform this application to three of the major questions in crowd work and gig work, identifying similarities and differences between historical piecework and modern on--demand work.

% \onlyinsubfile{\clearpage}
\subsection{A Piecework Primer}\label{sec:pieceworkPrimer} % ali adores alliteration
\topic{In this section we'll discuss the history of piecework at a high level
to contextualize the observations and conclusions we later make,
rather than provide a comprehensive historical account.}
% Instead, this section will attempt to provide a sense of orientation when thinking about piecework.
In other words, it will frame piecework,
in the early days of the Industrial Revolution \&
through the political and economic turmoil of the early and mid--\nth{20} century,
and into the \nth{21} century.
While the previous section provided a \textit{definition} of piecework,
this section attempts to shine a light on the \textit{zeitgeist} of piecework;
as close to a \textit{thick description} of piecework
as we can afford~\cite{geertz1973interpretation}.

@@

\topic{Piecework's popularity in the United States and Europe plummeted almost as quickly as it had climbed.}
\citeauthor{hart2013rise}'s work substantively explores the precipitous decline of piecework in the last third of the \nth{20} century.
In their work, \citeauthor{hart2013rise} offer a number of explanations for the sudden vanishing of piecework.
The salient suggestions include:
\begin{inlinelist}[itemjoin*={;~and~},itemjoin={;~}]
\item the emergence of more effective, more nuanced incentive models
--- rewarding teams for complex achievements, for instance
\item the shifting of these industries (manufacturing, clothing, etc\dots)
to other countries
\item the quality of ``multidimensional'' work becoming too difficult to evaluate
\end{inlinelist}~\cite{hart2013rise}.

% \onlyinsubfile{\clearpage}
\subsection{Case studies in piecework}
We'll discuss the context and factual details of three instructive cases of piecework to inform the three questions we investigate later ---
       % (\namerefl{sec:complexity}, \namerefl{sec:decomposition}, and \namerefl{sec:relationships}).}
\begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
  \item railroad and other industrial workers across the United States
  \item Airy's employment of \textit{human computers}
  \item domestic and farm work (in particular, the ``match--girls'').
\end{numberlist}
We'll approach this by looking at familiar cases of decomposed, distributed work (industrial workers),
and make our way backwards through the \nth{20} and then \nth{19} century,
tracing the origins of the ideas that

@@

% \ali{I'm looking into \citeauthor{grier2013computers}'s book to see if there's more about the relationships workers had with Airy.
% Maybe I can/should expand this to discuss other implementations of human computation around this time (if I find any good anecdotes).}
\onlyinsubfile{\clearpage}
\subsubsection{Railroad and other Industrial Workers}
{Piecework might be most familiar to the HCI researcher in the context of the assembly line, which largely defined manufacturing through the \nth{20} century.}
It was here that
scientific management, Fordism, and Taylorism
dramatically influenced how workers were managed and
the ways in which workers were perceived and envisioned paradigmatically;
it's here we'll start our overview of piecework's history through case studies.

\topic{Piecework through the \nth{20} century centered around auto and other mass manufacturing, but
found its way into the war effort during World War II.}
With the vast majority of men drafted into service,
factories found themselves turning to
a predominantly female workforce that had neither
the formal training nor
the years of apprenticeship experience
that conventional workers would have had.
Rather than attempting to train this new labor force in every aspect of industrial work,
these women were trained for individual tasks
% --- such as riveting component to another ---
and assigned to that task. %, decomposed from a more comprehensive task.
One might reflect on the observation that ``Rosie the Riveter'',
an icon of \nth{20} century America who
represented empowerment and opportunity for women~\cite{honey1985creating},
was a pieceworker~\cite{davies2014origins}.

% This approach would turn out to yield two benefits:
% \begin{numberlist}[itemjoin*={;~and~},itemjoin={;~}]
% \item these workers could be productive with significantly less ``on--boarding'' cost
% \item workers' performance could be tracked \& evaluated, and better methods could be disseminated to workers.
% \end{numberlist}

{With \citeauthor{taylor1914principles}'s formalization of scientific management in \textit{Taylorism}
(and Henry Ford's eponymously named \textit{Fordism}),
piecework in the early and mid--\nth{20} century surged, especially in industrial work.}
Scientific management promised that the careful measurement of workers would yield
higher efficiency and output~\cite{taylor1914principles,towardsGlobalFordism}.
While \citeauthor{Brown01041990} points out that
piecework dramatically advanced the instrumented measurement of workers,
in \citeauthor{taylor1914principles}'s time highly instrumented,
automatic measurement of workers was all but impossible~\cite{Brown01041990}.
Instead, managers conducted ``stop watch time studies''~\cite{nadworny1955scientific},
using completion times to inform per--task compensation.
% Under this regime, piecework became commonplace in railroad maintenance and a number 
% This approach led to the adoption of piecework systems in myriad industrial labor verticals, such as
% railroad maintenance, automobile manufacturing, and later weapons manufacturing.
% \ali{???? better word for ``weapons manufacturing''?}
% Instead, as a result,
% the distillation of work into smaller units ultimately
% bottomed out with tasks as small as could be usefully measured~\cite{10.2307/23702539}.

@@

\topic{The 1930s represented a boom for piecework on an unprecedented scale,
especially among engineering and metalworking industries.}
\citeauthor{hart2013rise} characterize the 1930s
--- and more broadly the first half of the \nth{20} century ---
as the ``heyday'' of the use of piecework.
They attribute this to the shortage of male workers,
who would have gone through a conventional apprenticeship process
affording them more comprehensive knowledge of the total scope of work.

\onlyinsubfile{\clearpage}
\subsubsection{Domestic and Farmhand Labor}
\topic{When piecework entered the American economy, it was not used for complex work.}
One reason for low complexity was workers' skills:
it was infeasible to provide new pieceworkers with the comprehensive education
that apprenticeships imparted~\cite{hart2013rise}.
So, initially piecework in the United States arose for farm work, and as
\citeauthor{hughRaynbirdTaskWork} and others discuss,
the practice remained relatively obscure until
it blossomed in the textile industry~\cite{hughRaynbirdTaskWork}.
The complexity of the work remained low at the turn of
the \nth{20} century as piecework saturated New York City~\cite{riisOtherSideLives}.
However, writers of the time focused their attention on
wage~\cite{burton1899commercial} and
management regimes~\cite{norton1900textile}
rather than training.
Mass manufacturing,
such as sewing garments~\cite{riisOtherSideLives} and making matchsticks~\cite{10.2307/3827491},
flourished under piecework systems in densely populated cities.

@@

\topic{The match--girls strike foreshadowed both collective action victories and an emerging paradigm regarding worker management.}
Coal miners won a nationwide individual minimum wage after The Great Coal Strike of 1912~\cite{10.2307/2221944}.
Garment workers in Philadelphia secured collective bargaining rights in 1915 after a prolonged strike and threat of a second~\cite{10.2307/41829256}.
It was in the midst of this time that \citeauthor{taylor1914principles} published
the work for which he would later be called ``the father of scientific management''~\cite{RePEc:mtp:titles:0262612062}.
It was this framing on work \& worker management that gave workers a concrete adversary
--- if not Taylor, then \textit{Taylorism}  ---
against which to rally~\cite{jacoby1983union,parker1920casual}.

\onlyinsubfile{\clearpage}
\subsubsection{Airy's Computers}
\topic{Some of the first systematic cases of what we would recognize as crowd work
can be found in the study of astronomy.}
In the \nth{19} century, the calculation of celestial bodies had become
a competitive field with 
Airy needed to compute tables that would
allow sailors to locate themselves by starlight from sea.
This work ostensibly called for educated people who comprehensively understood mathematics.
Airy realized that he could break the tasks down and delegate the constituent parts
to ``human computers'' who
``\dots~possessed the basic skills of mathematics,
including `Arithmetic, the use of Logarithms, and Elementary Algebra'~''~\cite{grier2013computers}.
As a result, many of Airy's computers had relatively rudimentary educations
compared to the background of education that typically worked in the calculation of solar tables.
Airy distributed tasks by mail,
allowing work to be completed by a somewhat geographically distributed workforce,
and paid for each piece of work completed.
% \ali{also talk about the challenge/competition Airy was in with other astronomers,
% and the approach of verifying worker results}
% Airy's method of distributing work was unorthodox, but his workforce management style was especially unconventional.
% For one thing,
Airy also instituted a policy of firing his computers once they reached age 23.

\ali{THIS GOES LATER: This practice ensured two outcomes that arguably disfavored workers.
\begin{Numberlist}
\item it eliminated any potential to advance professionally, as workers' careers in this area ended relatively early in their careers,
      and without formal education in mathematics they struggled to find work for which their experience was meaningful.
\item it limited workers' ability to organize
      by ensuring that workers hardly spent sufficient time to successfully rally their peers.
\end{Numberlist}}

\onlyinsubfile{
  % \printbibliography
  }
